---
title: "Bell Test: Symplectic Contextual Erasure"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)

data_dir <- "/Volumes/SanDisk4TB/SternBrocot-data/06_bell_test"

# Load the primary summary statistics
summary_dt <- fread(file.path(data_dir, "summary_stats.csv.gz"))

# Reference Functions
quantum_ref <- function(bob_angle, alice_fixed) {
  -cos(alice_fixed - bob_angle)
}

classical_ref <- function(bob_angle, alice_fixed) {
  rel <- (alice_fixed - bob_angle) %% (2 * pi)
  ifelse(rel <= pi, -1 + (2 * rel / pi), 3 - (2 * rel / pi))
}

# CHSH Interpolation
e1 <- approx(summary_dt$bob_angle_rad, summary_dt$correlation_alice_0,  xout = pi/4)$y
e2 <- approx(summary_dt$bob_angle_rad, summary_dt$correlation_alice_0,  xout = 3*pi/4)$y
e3 <- approx(summary_dt$bob_angle_rad, summary_dt$correlation_alice_90, xout = pi/4)$y
e4 <- approx(summary_dt$bob_angle_rad, summary_dt$correlation_alice_90, xout = 3*pi/4)$y
s_val <- abs(e1 - e2 + e3 + e4)

# --- Diagnostic Data Prep ---
all_files <- list.files(data_dir, pattern = "erasure_.*\\.csv\\.gz")
alice_files <- grep("alice", all_files, value = TRUE)
bob_files <- grep("bob", all_files, value = TRUE)

parse_angle <- function(file_names) {
  as.numeric(gsub("erasure_[^_]+_([-0-9.]+)\\.csv\\.gz", "\\1", file_names))
}

# Load Alice 0 data for internal diagnostic plots
alice_numeric_angles <- parse_angle(alice_files)
file_key_0 <- alice_files[order(alice_numeric_angles)[1]]
alice_0_data <- fread(file.path(data_dir, file_key_0))

# Calculate average algorithmic complexity and found rate across Bob's sweeps
complexity_dt <- rbindlist(lapply(bob_files, function(f) {
  dt <- fread(file.path(data_dir, f), select = c("program_length", "found"))
  data.table(
    bob_angle_rad = parse_angle(f),
    avg_length = mean(dt$program_length, na.rm = TRUE),
    found_rate = mean(dt$found, na.rm = TRUE) * 100
  )
}))

# Calculate Alice's baseline found rate for the summary table
alice_found_rate <- mean(alice_0_data$found, na.rm = TRUE) * 100
```

## Statistical Summary

## Statistical Summary

| Metric | Value |
| :--- | :--- |
| **CHSH S-Value** | **`r round(s_val, 5)`** |
| **Classical Limit** | 2.0 |
| **Quantum Limit** | 2.828 |
| **Algorithm Found Rate** | **`r round(alice_found_rate, 2)`%** |

---

## Alice Fixed at 0°

Correlation $E(0, \theta_B)$ compared against the quantum $-\cos(\theta)$ baseline (dashed) and the linear classical baseline (solid).

```{r plot-alice-0, echo=FALSE, fig.width=10, fig.height=5}
ggplot(summary_dt, aes(x = bob_angle_rad %% (2*pi))) +
  stat_function(fun = function(x) classical_ref(x, 0), color = "gray85", linewidth = 1.2) +
  stat_function(fun = function(x) quantum_ref(x, 0), color = "gray50", linetype = "dashed", linewidth = 0.6) +
  geom_point(aes(y = correlation_alice_0), color = "black", size = 0.5) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/4), 
                     labels = c("0", "π/4", "π/2", "3π/4", "π", "5π/4", "3π/2", "7π/4", "2π")) +
  ylim(-1.1, 1.1) +
  labs(title = "Detector Correlation (Alice Fixed at 0°)", x = "Bob Angle (rad)", y = "Correlation E") +
  theme_minimal()
```

---

## Alice Fixed at 90° (π/2)

Correlation plotted against the Phase Difference $(\theta_B - \theta_A)$.

```{r plot-alice-90, echo=FALSE, fig.width=10, fig.height=5}
ggplot(summary_dt, aes(x = phase_diff_90)) +
  stat_function(fun = function(x) classical_ref(x, 0), color = "gray85", linewidth = 1.2) +
  stat_function(fun = function(x) quantum_ref(x, 0), color = "gray50", linetype = "dashed", linewidth = 0.6) +
  geom_point(aes(y = correlation_alice_90), color = "black", size = 0.5) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/4), 
                     labels = c("0", "π/4", "π/2", "3π/4", "π", "5π/4", "3π/2", "7π/4", "2π")) +
  ylim(-1.1, 1.1) +
  labs(title = "Detector Correlation (Alice Fixed at 90°)", 
       x = "Phase Difference (θ_B - θ_A) (rad)", 
       y = "Correlation E") +
  theme_minimal()
```

### CHSH Breakdown
```text
E(0,  pi/4):  `r sprintf("%7.4f", e1)`
E(0, 3pi/4):  `r sprintf("%7.4f", e2)`
E(90, pi/4):  `r sprintf("%7.4f", e3)`
E(90, 3pi/4): `r sprintf("%7.4f", e4)`
-----------------------
S-Value:      `r sprintf("%7.4f", s_val)`
```

---

## Erasure Transfer Function

This plot visualizes the internal mathematical "warp" of the contextual erasure. It shows how the continuous local hidden variable (microstate) is shifted into its final macroscopic physical distance.

```{r plot-transfer-function, echo=FALSE, fig.width=10, fig.height=5}
ggplot(alice_0_data, aes(x = microstate, y = erasure_distance)) +
  geom_point(color = "dodgerblue", size = 0.5, alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/2), 
                     labels = c("0", "π/2", "π", "3π/2", "2π")) +
  labs(title = "Microstate vs. Erasure Distance", 
       x = "Initial Microstate (rad)", 
       y = "Erasure Distance Shift") +
  theme_minimal()
```

---

## Erasure Distance vs. Uncertainty

This plot visualizes the relationship between the physical shift required to snap to a rational macrostate (`erasure_distance`) and the fundamental uncertainty of that resulting state. It reveals the precision trade-offs inherent in the contextual erasure process.

```{r plot-distance-vs-uncertainty, echo=FALSE, fig.width=10, fig.height=5}
ggplot(alice_0_data, aes(x = uncertainty, y = erasure_distance)) +
  geom_point(color = "darkorange", size = 0.5, alpha = 0.5) +
  labs(title = "Erasure Distance vs. Uncertainty", 
       x = "Uncertainty", 
       y = "Erasure Distance") +
  theme_minimal()
```

---

## Algorithmic Complexity of Measurement

This plot reveals the thermodynamic cost of the measurement context. As the particle's relative angle changes, the computational effort (program length) required to map the microstate to a rational Stern-Brocot macrostate fluctuates in a fractal pattern.

```{r plot-complexity, echo=FALSE, fig.width=10, fig.height=5}
# Plot the raw integer program lengths against the microstate for a single detector
ggplot(alice_0_data, aes(x = microstate, y = program_length)) +
  geom_point(color = "darkred", size = 1, alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/2), 
                     labels = c("0", "π/2", "π", "3π/2", "2π")) +
  labs(title = "Stern-Brocot Program Length vs Particle Microstate", 
       x = "Initial Microstate (rad)", 
       y = "Program Length (Steps)") +
  theme_minimal()
```

---

## Erasure Success Rate

This plot verifies the percentage of microstates successfully resolved to a rational macrostate before hitting the maximum search depth limit (`max_depth = 2000`). For the simulation to be valid, this must remain at 100%.

```{r plot-found-rate, echo=FALSE, fig.width=10, fig.height=3}
ggplot(complexity_dt, aes(x = bob_angle_rad, y = found_rate)) +
  geom_line(color = "forestgreen", linewidth = 1.2) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/2), 
                     labels = c("0", "π/2", "π", "3π/2", "2π")) +
  scale_y_continuous(limits = c(0, 105), breaks = c(0, 25, 50, 75, 100)) +
  labs(title = "Algorithm Found Rate vs Detector Angle", 
       x = "Bob Angle (rad)", 
       y = "Found Rate (%)") +
  theme_minimal()
```

---

## Quantization into Rational Macrostates

This histogram demonstrates how the contextual erasure acts as a fractal attractor, quantizing the uniform continuous microstates into discrete rational fractions on the Stern-Brocot tree.

```{r plot-attractors, echo=FALSE, fig.width=10, fig.height=5}
ggplot(alice_0_data, aes(x = macrostate)) +
  geom_histogram(bins = 200, fill = "purple", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Resolved Macrostates", 
       x = "Macrostate Value", 
       y = "Frequency") +
  theme_minimal()
```

## Macrostate vs. Initial Microstate

This plot maps the initial continuous hidden variable (microstate) directly to the resolved Stern-Brocot fraction (macrostate). In a purely classical model, this forms a rigid "Devil's Staircase" along a linear diagonal.

```{r plot-macro-vs-micro, echo=FALSE, fig.width=10, fig.height=5}
ggplot(alice_0_data, aes(x = microstate, y = macrostate)) +
  geom_point(color = "darkgreen", size = 0.5, alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/2), 
                     labels = c("0", "π/2", "π", "3π/2", "2π")) +
  labs(title = "Resolved Macrostate vs Initial Microstate", 
       x = "Initial Microstate (rad)", 
       y = "Quantized Macrostate (Normalized)") +
  theme_minimal()
```
