---
title: "Bell Test: Symplectic Contextual Erasure (Non-Local Dynamic)"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)

# Ensure this matches exactly with the generation script
data_dir <- "/Volumes/SanDisk4TB/SternBrocot-data/06_bell_test"

# Load the primary summary statistics
summary_dt <- fread(file.path(data_dir, "summary_stats.csv.gz"))

# Split into Alice 0 and Alice 90 datasets for the plots
alice_0_dt <- summary_dt[abs(alice_angle_rad - 0.0) < 1e-4]
alice_90_dt <- summary_dt[abs(alice_angle_rad - (pi/2)) < 1e-4]
alice_90_dt[, phase_diff_90 := (bob_angle_rad - pi/2) %% (2 * pi)]

# Reference Functions
quantum_ref <- function(bob_angle, alice_fixed) {
  -cos(alice_fixed - bob_angle)
}

classical_ref <- function(bob_angle, alice_fixed) {
  rel <- (alice_fixed - bob_angle) %% (2 * pi)
  ifelse(rel <= pi, -1 + (2 * rel / pi), 3 - (2 * rel / pi))
}

# CHSH Interpolation
e1 <- approx(alice_0_dt$bob_angle_rad, alice_0_dt$correlation,  xout = pi/4)$y
e2 <- approx(alice_0_dt$bob_angle_rad, alice_0_dt$correlation,  xout = 3*pi/4)$y
e3 <- approx(alice_90_dt$bob_angle_rad, alice_90_dt$correlation, xout = pi/4)$y
e4 <- approx(alice_90_dt$bob_angle_rad, alice_90_dt$correlation, xout = 3*pi/4)$y
s_val <- abs(e1 - e2 + e3 + e4)

# --- Diagnostic Data Prep ---
all_files <- list.files(data_dir, pattern = "bell_pair_.*\\.csv\\.gz")

# Parse both alpha and beta dynamically
parse_alpha_angle <- function(file_names) {
  as.numeric(gsub(".*_alpha_([-0-9.]+)_beta.*", "\\1", file_names))
}
parse_beta_angle <- function(file_names) {
  as.numeric(gsub(".*_beta_([-0-9.]+)\\.csv\\.gz", "\\1", file_names))
}

all_alphas <- parse_alpha_angle(all_files)

# 1. Load data for Alpha ~ 0
alice_0_files <- all_files[abs(all_alphas - 0.0) < 1e-4]
beta_angles_0 <- parse_beta_angle(alice_0_files)
file_key_0 <- alice_0_files[order(abs(beta_angles_0))[1]] # File where Bob is near 0
raw_0 <- fread(file.path(data_dir, file_key_0))
raw_0[, alpha_label := "Alpha = 0 Radians (0°)"]

# 2. Load data for Alpha ~ pi/2
alice_90_files <- all_files[abs(all_alphas - (pi/2)) < 1e-4]
beta_angles_90 <- parse_beta_angle(alice_90_files)
file_key_90 <- alice_90_files[order(abs(beta_angles_90))[1]] # File where Bob is near 0
raw_90 <- fread(file.path(data_dir, file_key_90))
raw_90[, alpha_label := "Alpha = π/2 Radians (90°)"]

# Combine diagnostic files
raw_diagnostic_data <- rbindlist(list(raw_0, raw_90))

# Map the strictly named 28-column paired format to the original ggplot variables
erasure_data <- data.table(
  microstate = raw_diagnostic_data$microstate,
  erasure_distance = raw_diagnostic_data$bob_erasure_distance,
  minimal_action_state = raw_diagnostic_data$bob_minimal_action_state,
  uncertainty = raw_diagnostic_data$delta_theta,
  program_length = raw_diagnostic_data$bob_program_length,
  found = raw_diagnostic_data$bob_found,
  alpha_label = raw_diagnostic_data$alpha_label
)

# Calculate average algorithmic complexity, found rate, AND extract delta_theta across Bob's sweeps
complexity_dt_0 <- rbindlist(lapply(alice_0_files, function(f) {
  dt <- fread(file.path(data_dir, f), select = c("bob_program_length", "bob_found", "delta_theta"))
  data.table(
    bob_angle_rad = parse_beta_angle(f),
    avg_length = mean(dt$bob_program_length, na.rm = TRUE),
    found_rate = mean(dt$bob_found, na.rm = TRUE) * 100,
    delta_theta = dt$delta_theta[1], # Grab the shared boundary for this specific context
    alpha_label = "Alpha = 0 Radians (0°)"
  )
}))

complexity_dt_90 <- rbindlist(lapply(alice_90_files, function(f) {
  dt <- fread(file.path(data_dir, f), select = c("bob_program_length", "bob_found", "delta_theta"))
  data.table(
    bob_angle_rad = parse_beta_angle(f),
    avg_length = mean(dt$bob_program_length, na.rm = TRUE),
    found_rate = mean(dt$bob_found, na.rm = TRUE) * 100,
    delta_theta = dt$delta_theta[1], # Grab the shared boundary for this specific context
    alpha_label = "Alpha = π/2 Radians (90°)"
  )
}))

complexity_dt <- rbindlist(list(complexity_dt_0, complexity_dt_90))

# Calculate baseline found rate for the summary table
overall_found_rate <- mean(erasure_data$found, na.rm = TRUE) * 100
```

## Statistical Summary

| Metric | Value |
| :--- | :--- |
| **CHSH S-Value** | **`r round(s_val, 5)`** |
| **Classical Limit** | 2.0 |
| **Quantum Limit** | 2.828 |
| **Algorithm Found Rate** | **`r round(overall_found_rate, 2)`%** |

---

## Alice Fixed at 0°

Correlation $E(0, \theta_B)$ compared against the quantum $-\cos(\theta)$ baseline (dashed) and the linear classical baseline (solid).

```{r plot-alice-0, echo=FALSE, fig.width=10, fig.height=5}
ggplot(alice_0_dt, aes(x = bob_angle_rad %% (2*pi))) +
  stat_function(fun = function(x) classical_ref(x, 0), color = "gray85", linewidth = 1.2) +
  stat_function(fun = function(x) quantum_ref(x, 0), color = "gray50", linetype = "dashed", linewidth = 0.6) +
  geom_point(aes(y = correlation), color = "black", size = 0.5) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/4), 
                     labels = c("0", "π/4", "π/2", "3π/4", "π", "5π/4", "3π/2", "7π/4", "2π")) +
  ylim(-1.1, 1.1) +
  labs(title = "Detector Correlation (Alice Fixed at 0°)", x = "Bob Angle (rad)", y = "Correlation E") +
  theme_minimal()
```

---

## Alice Fixed at 90° (π/2)

Correlation plotted against the Phase Difference $(\theta_B - \theta_A)$.

```{r plot-alice-90, echo=FALSE, fig.width=10, fig.height=5}
ggplot(alice_90_dt, aes(x = phase_diff_90)) +
  stat_function(fun = function(x) classical_ref(x, 0), color = "gray85", linewidth = 1.2) +
  stat_function(fun = function(x) quantum_ref(x, 0), color = "gray50", linetype = "dashed", linewidth = 0.6) +
  geom_point(aes(y = correlation), color = "black", size = 0.5) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/4), 
                     labels = c("0", "π/4", "π/2", "3π/4", "π", "5π/4", "3π/2", "7π/4", "2π")) +
  ylim(-1.1, 1.1) +
  labs(title = "Detector Correlation (Alice Fixed at 90°)", 
       x = "Phase Difference (θ_B - θ_A) (rad)", 
       y = "Correlation E") +
  theme_minimal()
```

### CHSH Breakdown

```text
E(0,  pi/4):  `r sprintf("%7.4f", e1)`
E(0, 3pi/4):  `r sprintf("%7.4f", e2)`
E(90, pi/4):  `r sprintf("%7.4f", e3)`
E(90, 3pi/4): `r sprintf("%7.4f", e4)`
-----------------------
S-Value:      `r sprintf("%7.4f", s_val)`
```

---

## Erasure Transfer Function

This plot visualizes the internal mathematical "warp" of the contextual erasure. It shows how the continuous local hidden variable (microstate) is shifted into its final macroscopic physical distance.

```{r plot-transfer-function, echo=FALSE, fig.width=10, fig.height=5}
ggplot(erasure_data, aes(x = microstate, y = erasure_distance)) +
  geom_point(color = "dodgerblue", size = 0.5, alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/2), 
                     labels = c("0", "π/2", "π", "3π/2", "2π")) +
  facet_wrap(~ alpha_label) +
  labs(title = "Microstate vs. Erasure Distance", 
       subtitle = "Diagnostic data showing Bob's internal state (Bob near 0 rad)",
       x = "Initial Microstate (rad)", 
       y = "Erasure Distance Shift") +
  theme_minimal()
```

---

## Erasure Distance vs. Uncertainty

This plot visualizes the relationship between the physical shift required to snap to a rational minimal_action_state (`erasure_distance`) and the fundamental uncertainty of that resulting state. It reveals the precision trade-offs inherent in the contextual erasure process.

```{r plot-distance-vs-uncertainty, echo=FALSE, fig.width=10, fig.height=5}
ggplot(erasure_data, aes(x = uncertainty, y = erasure_distance)) +
  geom_point(color = "darkorange", size = 0.5, alpha = 0.5) +
  facet_wrap(~ alpha_label) +
  labs(title = "Erasure Distance vs. Shared Non-Local Tolerance", 
       subtitle = "Diagnostic data showing Bob's internal state (Bob near 0 rad)",
       x = "Shared Tolerance (Uncertainty)", 
       y = "Erasure Distance") +
  theme_minimal()
```

---

## Algorithmic Complexity of Measurement

This plot reveals the thermodynamic cost of the measurement context. As the particle's relative angle changes, the computational effort (program length) required to map the microstate to a rational Stern-Brocot minimal_action_state fluctuates in a fractal pattern.

```{r plot-complexity, echo=FALSE, fig.width=10, fig.height=5}
ggplot(erasure_data, aes(x = microstate, y = program_length)) +
  geom_point(color = "darkred", size = 1, alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/2), 
                     labels = c("0", "π/2", "π", "3π/2", "2π")) +
  facet_wrap(~ alpha_label) +
  labs(title = "Stern-Brocot Program Length vs Particle Microstate", 
       subtitle = "Diagnostic data showing Bob's internal state (Bob near 0 rad)",
       x = "Initial Microstate (rad)", 
       y = "Program Length (Steps)") +
  theme_minimal()
```

---

## Erasure Success Rate

This plot verifies the percentage of microstates successfully resolved to a rational minimal_action_state before hitting the maximum search depth limit (`max_depth = 2000`). For the simulation to be valid, this must remain at 100%.

```{r plot-found-rate, echo=FALSE, fig.width=10, fig.height=5}
ggplot(complexity_dt, aes(x = bob_angle_rad, y = found_rate)) +
  geom_line(color = "forestgreen", linewidth = 1.2) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/2), 
                     labels = c("0", "π/2", "π", "3π/2", "2π")) +
  scale_y_continuous(limits = c(0, 105), breaks = c(0, 25, 50, 75, 100)) +
  facet_wrap(~ alpha_label) +
  labs(title = "Algorithm Found Rate vs Detector Angle", 
       x = "Bob Angle (rad)", 
       y = "Found Rate (%)") +
  theme_minimal()
```

---

## Quantization into Rational minimal_action_states

This histogram demonstrates how the contextual erasure acts as a fractal attractor, quantizing the uniform continuous microstates into discrete rational fractions on the Stern-Brocot tree.

```{r plot-attractors, echo=FALSE, fig.width=10, fig.height=5}
ggplot(erasure_data, aes(x = minimal_action_state)) +
  geom_histogram(bins = 200, fill = "purple", color = "black", alpha = 0.7) +
  facet_wrap(~ alpha_label) +
  labs(title = "Distribution of Resolved minimal_action_states", 
       subtitle = "Diagnostic data showing Bob's internal state (Bob near 0 rad)",
       x = "minimal_action_state Value", 
       y = "Frequency") +
  theme_minimal()
```

## minimal_action_state vs. Initial Microstate

This plot maps the initial continuous hidden variable (microstate) directly to the resolved Stern-Brocot fraction (minimal_action_state).

```{r plot-macro-vs-micro, echo=FALSE, fig.width=10, fig.height=5}
ggplot(erasure_data, aes(x = microstate, y = minimal_action_state)) +
  geom_point(color = "darkgreen", size = 0.5, alpha = 0.5) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/2), 
                     labels = c("0", "π/2", "π", "3π/2", "2π")) +
  facet_wrap(~ alpha_label) +
  labs(title = "Resolved minimal_action_state vs Initial Microstate", 
       subtitle = "Diagnostic data showing Bob's internal state (Bob near 0 rad)",
       x = "Initial Microstate (rad)", 
       y = "Quantized minimal_action_state (Normalized)") +
  theme_minimal()
```

---

## Non-Local Action Boundary vs Detector Angle

This plot maps the shared thermodynamic boundary (`delta_theta`) as Bob rotates his measurement apparatus, demonstrating how the potential action limits pulse relative to the fixed context of Alice's magnet.

```{r plot-delta-theta, echo=FALSE, fig.width=10, fig.height=5}
ggplot(complexity_dt, aes(x = bob_angle_rad, y = delta_theta)) +
  geom_line(color = "blue", linewidth = 1.2) +
  scale_x_continuous(breaks = seq(0, 2*pi, by = pi/2), 
                     labels = c("0", "π/2", "π", "3π/2", "2π")) +
  facet_wrap(~ alpha_label) +
  labs(title = "Shared Halting Tolerance (delta_theta) vs Bob Angle",
       x = "Bob Angle (rad)",
       y = "Delta Theta (Tolerance)") +
  theme_minimal()
```
